---
title: "R Assignment"
output: html_notebook
---
Setting the directory
```{r}
setwd("~/Desktop/BioDataSkillsClass/UNIX_Assignment")
```

Loading files to R

```{r}
fang <- read.table("fang_et_al_genotypes.txt", header = TRUE )
position <- read.table("snp_position.txt", header = TRUE, fill = TRUE)
```
#Part 1

##Data Inspection
```{r}
str(fang)
str(position)
dim(fang)
dim(position)
nrow(fang)
nrow(position)
ncol(fang)
ncol(position)

head(fang, n = 1)
head(position, n = 3)
```

##Data processing

Subset necessary columns
```{r}
fang_subset <- subset(fang, select = -c(1:2))
```

Extracting maize and teosinte data
```{r}
maize <- c("ZMMIL","ZMMLR","ZMMMR")
maize_extracted <- fang_subset[fang_subset$Group %in% maize,]

teosinte <- c("ZMPBA", "ZMPIL", "ZMPJA")
teosinte_extracted <- fang_subset[fang_subset$Group %in% teosinte,]
```

Transpose the data
```{r}
maize_transposed <- as.data.frame(t(maize_extracted[,-1]))

teosinte_transposed <- as.data.frame(t(teosinte_extracted[,-1]))
```

Subset columns of snp_position file that matter
```{r}
position_subset <- subset(position, select = c(1,3,4))
```

Sort files
```{r}
maize_sorted <- maize_transposed[order(row.names(maize_transposed)),]
teosinte_sorted <- teosinte_transposed[order(row.names(teosinte_transposed)),]
position_sorted <- position_subset[order(position_subset$SNP_ID),]
```

Remove data from "unknown and random" chromosomes in SNP position file
```{r}
chrom <- c(1,2,3,4,5,6,7,8,9,10)
position_sorted_selected <- subset(position_sorted, Chromosome %in% chrom)
position_sorted <- position_sorted_selected[-1,]
```
Making row names as columns so I can join the data frames
```{r}
maize_row <- add_rownames(maize_sorted, "SNP_ID")
teosinte_row <- add_rownames(teosinte_sorted, "SNP_ID")
```

Merge maize and teosinte data with snp position data
```{r}
maize_merged <- merge(position_sorted, maize_row, by="SNP_ID")
teosinte_merged <- merge(position_sorted, teosinte_row, by="SNP_ID")
``` 

Sort by SNP position - ascending order
```{r}
maize_sorted_by_position <- arrange(maize_merged, Position)  #ascending position
teosinte_sorted_by_position <- arrange(teosinte_merged, Position) #ascending position 
```

Substitute symbols - no data by ?
```{r}
gsub("","?", maize_sorted_by_position) 
gsub("", "?", teosinte_sorted_by_position)
```

Sort by SNP position - descending order
```{r}
maize_desc_sorted_by_position <- arrange(maize_merged, desc(Position))
teosinte_desc_sorted_by_position <- arrange(teosinte_merged, desc(Position))
```
Substitute symbols - no data by -
 
```{r}
gsub("","-", maize_desc_sorted_by_position) 
gsub("","-", teosinte_desc_sorted_by_position)
```

Extract data by chromosome number
```{r}
for(i in 1:10){
 temp_df<-maize_sorted_by_position[maize_sorted_by_position[,2]==i,]
 write.csv(temp_df, paste("maize_chr",i,".csv",sep=""),row.names=F)}

for(i in 1:10){
 temp_df<-teosinte_sorted_by_position[teosinte_sorted_by_position[,2]==i,]
 write.csv(temp_df, paste("teosinte_chr",i,".csv",sep=""),row.names=F)}

for(i in 1:10){
 temp_df<-maize_desc_sorted_by_position[maize_desc_sorted_by_position[,2]==i,]
 write.csv(temp_df, paste("maize_desc_chr",i,".csv",sep=""),row.names=F)}

for(i in 1:10){
 temp_df<-teosinte_desc_sorted_by_position[teosinte_desc_sorted_by_position[,2]==i,]
 write.csv(temp_df, paste("teosinte_desc_chr",i,".csv",sep=""),row.names=F)}
```


###Part 1 finished

#Part 2

Tidying the data 
```{r}
library(reshape2)
maize_reshaped <- melt(maize_merged, id=(c("SNP_ID","Chromosome","Position")))
teosinte_reshaped <- melt(teosinte_merged, id=(c("SNP_ID","Chromosome","Position")))
```

Plot total number of SNPs by chromosome number
```{r}

```

Create new column indicating if site is homozygous or heterozygous
```{r}

```

Substitute missing data with NA
```{r}

```

Sort dataframe by Species_ID and Group values
```{r}

```

Graph that shows the proportion of homozygous and heterozygous sites as well as missing data in each species
```{r}

```

